---
title: "Project 2"
author: "Jingcong Jiang and Haoran Gu"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(coin)
```

### part1
### First plot the data in order to get a first impression
```{r}
bacterial <- read.csv("bacterial.csv")

table(bacterial$Method)
table(bacterial$Method)/sum(table(bacterial$Method))

boxplot(Bacterial.Counts ~ Method, data = bacterial, main = "Bacterial colonies by Washing method", horizontal = TRUE)
ggplot(bacterial, aes(x = Bacterial.Counts)) + geom_histogram(binwidth = 8,color = "black",fill = "white") + facet_grid(Method ~.) +ggtitle("Bacterial colonies by Washing method")
```

### Use Kruskal-Wallis (K-W) Test and Wilcoxon Rank Sum test to do compare the methods
### First create a table which contains general information of the data
```{r}
bacterial$Rank = rank(bacterial$Bacterial.Counts, ties = "average")
Group.order = aggregate(Bacterial.Counts ~ Method, data = bacterial, mean)$Method
Xi = aggregate(Bacterial.Counts ~ Method, data = bacterial, mean)$Bacterial.Counts
si = aggregate(Bacterial.Counts ~ Method, data = bacterial, sd)$Bacterial.Counts
Ri = aggregate(Rank ~ Method, data = bacterial, sd)$Rank
ni = aggregate(Bacterial.Counts ~ Method, data = bacterial, length)$Bacterial.Counts
results = rbind(Xi,si,Ri,ni)
rownames(results) = c("Bacterial count Mean","Bacterical count SD","Rank Mean","Sample Size")
colnames(results) = as.character(Group.order)
results
```
### The result shows that alcohol spray is the most effective method to eliminate bacterial. But the effectiveness of the other three methods is not clear.

```{r}
#Antibacterial Soap vs Soap
new.data1 = bacterial[c(3,7,16,23,2,6,10,12,15,19,22,26),]
new.data1$Rank = rank(new.data1$Bacterial.Counts, ties = "average")
wilcox_test(Bacterial.Counts ~ as.factor(Method),new.data1, distribution = "exact",alternative = "less")
```
### We use alpha = 0.05 for the test
### P-value > 0.05, this suggests that the differences between Antibacterial Soap and Soap method are not clear.

```{r}
#Antibacterial Soap vs water
new.data2 = bacterial[c(3,7,16,23,1,5,9,14,18,21,25),]
new.data2$Rank = rank(new.data2$Bacterial.Counts, ties = "average")
wilcox_test(Bacterial.Counts ~ as.factor(Method),new.data2, distribution = "exact",alternative = "less")
```
### P-value < 0.05, this suggests that the bacterial count left using Antibacterial Soap is less than the bacterial count left using water.

```{r}
#Soap vs water
new.data3 = bacterial[c(2,6,10,12,15,19,22,26,1,5,9,14,18,21,25),]
new.data3$Rank = rank(new.data3$Bacterial.Counts, ties = "average")
wilcox_test(Bacterial.Counts ~ as.factor(Method),new.data3, distribution = "exact",alternative = "less")
```
### P-value > 0.05, this suggests that the differences between Soap and Water method are not clear.

### Even though effectiveness between Antibacterial Soap vs Soap method, and Soap vs Water method is not clear, the result of the contrast between Antibacterial Soap and water gives us a hint. It suggests that Antibacterial Soap is somehow the best washing method among these three methods, while Water is somehow the worst.















## Part 2
```{r}
mind.data = read.csv("Mind.csv")
mind.table = table(mind.data)
mind.table
```
```{r}
summary(mind.data)
```

### State the null and alternative hypothesis first

#### Ho: Treatment and Improvement are independent for patients

#### Ha: Treatment and Improvement are dependent for patients
```{r}
mind.test = chisq.test(mind.data$Treatment, mind.data$Improve, correct = FALSE)
mind.test
```

```{r}
eij = mind.test$expected

eij
```

```{r}
chi.sq.obs = as.numeric(mind.test$statistic)

chi.sq.obs
```
### Do permutations and calculate the permutation-based p value, here use R = 5000 

```{r}
chisq_test(Treatment ~ Improve, mind.data, distribution = approximate(nresample = 5000))
```
#### Use alpha = 0.01, our pvalue is less than alpha. So we reject Ho. We are 99% confident that Treatment and Improvement are dependent for patients.

### Do pair-wise comparison to check which specific subcategories are dependent
```{r}
n = sum(mind.table)
ni. = rowSums(mind.table)
n.j = colSums(mind.table)
all.pjG1 = mind.table[1,]/ni.[1] #all conditional probabilites for row 1
all.pjG2= mind.table[2,]/ni.[2] #all conditional probabilites for row 2
all.pbar = n.j/n #all probabilities regardless of group
all.Zij = c(all.pjG1 - all.pjG2)/sqrt(all.pbar*(1-all.pbar)*(1/ni.[1] + 1/ni.[2])) #The z-test-statistics
```

```{r}
R = 5000
r.perms.cutoff = sapply(1:R,function(i){
  perm.data = mind.data
  perm.data$Improve = sample(perm.data$Improve,nrow(perm.data),replace = FALSE)
  row.sum = rowSums(table(perm.data))
  col.sum = colSums(table(perm.data))
  all.pji = table(perm.data)[1,]/row.sum[1]
  all.pji.= table(perm.data)[2,]/row.sum[2]
  all.pbar = col.sum/sum(row.sum)
  all.Zij = c(all.pji - all.pji.)/sqrt(all.pbar*(1-all.pbar)*(1/row.sum[1] + 1/row.sum[2]))
  Q.r = max(abs(all.Zij))
  return(Q.r)
})
alpha = 0.01
cutoff.q = as.numeric(quantile(r.perms.cutoff,(1-alpha)))
cutoff.q
```

```{r}
all.Zij = matrix(all.Zij,nrow=  1)
colnames(all.Zij) = c("Major","Mild","Moderate","None")
rownames(all.Zij) = c("Medication vs. Therapy")
all.Zij
```
#### As the cutoff is 2.934, this suggests that Therapy as a treatment has more major improvement on patients.

#### According to the sign of the Z score, Medication treatment has more mild and none improvement, and therapy treatment has more major and moderate treatment.
